---
title: "Benchmarking ChEA3 API"
output: html_notebook
---

Load all ChEA3 API benchmarking results

```{r}

rm(list = ls())

library(ggpubr)


borda = read.table("/Volumes/backup2/ChEA3_API_Benchmark/meanrank_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)

toprank = read.table("/Volumes/backup2/ChEA3_API_Benchmark/toprank_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)

chea3_libs = read.table("/Volumes/backup2/ChEA3_API_Benchmark/library_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)



```

```{r}
library(goftest)

borda_ks = ad.test(borda[borda$class == 1, "scaled_rank"])
borda_ks

toprank_ks = ad.test(toprank[toprank$class == 1, "scaled_rank"])
toprank_ks

chea3_libs_ks = plyr::dlply(chea3_libs,plyr::.(library),function(x){
  return(ad.test(x[x$class == 1, "scaled_rank"]))
})

chea3_libs_ks[["Literature--ChIP-seq"]]

borda_ks_df = data.frame(Method = "MeanRank", 
  Pval = signif(-log(borda_ks$p.value),4),stringsAsFactors = F)

toprank_ks_df = data.frame(Method = "TopRank", Pval = signif(-log(toprank_ks$p.value),4), stringsAsFactors = F)

libs_ks_df = plyr::ldply(chea3_libs_ks,function(x){
  return(data.frame(Pval = signif(-log(x$p.value),4), stringsAsFactors = F))})





```



Compute bootstrapped ROC Curves

```{r}
library(ggplot2)

n=5000

bootstrap_ROCcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    roc = PRROC::roc.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(roc$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + roc$auc
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(FPR = mean(x[,1]), TPR = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
}

#chea3 borda
borda_roc = bootstrap_ROCcurve(1-borda$scaled_rank[borda$class==1],1-borda$scaled_rank[borda$class==0],n)

#chea3 toprank
toprank_roc = bootstrap_ROCcurve(1-toprank$scaled_rank[toprank$class==1],1-toprank$scaled_rank[toprank$class==0], n)


#chea3 libs
chea3_libs_roc = plyr::dlply(chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_ROCcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot Bootstrapped ROC curves
```{r}
roc = data.frame()

roc = rbind(roc, data.frame(Method = "MeanRank",
  FPR = borda_roc$approx$x, TPR = borda_roc$approx$y, stringsAsFactors = F))

roc = rbind(roc, data.frame(Method = "TopRank",
  FPR = toprank_roc$approx$x, TPR = toprank_roc$approx$y, stringsAsFactors = F))

libs_roc = plyr::ldply(chea3_libs_roc,function(lib){
  return(data.frame(
    FPR = lib$approx$x, TPR = lib$approx$y, stringsAsFactors = F))
})

libs_roc$Method = gsub("--"," ",libs_roc$library)
libs_roc$library = NULL

roc = rbind(roc, libs_roc)

roc = roc[order(roc$Method),]
bootstrap_roc_plot = ggline(roc, x = "FPR", y = "TPR", color = "Method", plot_type = "l",numeric.x.axis = T) 

bootstrap_roc_plot

```

Compute bootstrapped PR curves

```{r}

n=5000

bootstrap_PRcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    pr = PRROC::pr.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(pr$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + pr$auc.integral
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(recall = mean(x[,1]), precision = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
 }


#chea3 borda
borda_pr = bootstrap_PRcurve(1-borda$scaled_rank[borda$class==1],1-borda$scaled_rank[borda$class==0],n)

#chea3 toprank
toprank_pr = bootstrap_PRcurve(1-toprank$scaled_rank[toprank$class==1],1-toprank$scaled_rank[toprank$class==0], n)


#chea3 libs
chea3_libs_pr = plyr::dlply(chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_PRcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot bootstrapped PR curves

```{r}

pr = data.frame()

pr = rbind(pr, data.frame(Method = "MeanRank",
  Recall = borda_pr$approx$x, Precision = borda_pr$approx$y, stringsAsFactors = F))

pr = rbind(pr, data.frame(Method = "TopRank",
  Recall = toprank_pr$approx$x, Precision = toprank_pr$approx$y, stringsAsFactors = F))

libs_pr = plyr::ldply(chea3_libs_pr,function(lib){
  return(data.frame(
    Recall = lib$approx$x, Precision = lib$approx$y, stringsAsFactors = F))
})

libs_pr$Method = gsub("--"," ",libs_pr$library)
libs_pr$library = NULL

pr = rbind(pr, libs_pr)

pr = pr[order(pr$Method),]
bootstrap_pr_plot = ggline(pr, x = "Recall", y = "Precision", color = "Method", plot_type = "l",numeric.x.axis = T) 

bootstrap_pr_plot

```


Compute and plot bridge plots 
```{r}


borda_ecdf = ecdf(borda$scaled_rank[borda$class == 1])

toprank_ecdf = ecdf(toprank$scaled_rank[toprank$class == 1])

chea3_libs_ecdf = plyr::dlply(chea3_libs,plyr::.(library),function(lib){
  return(ecdf(lib$scaled_rank[lib$class == 1]))
})

rand_ecdf = ecdf(runif(sum(borda$class == 1), min = 0, max = 1))

x = seq(0,1,0.001)

bridge_df = data.frame()

borda_bridge = data.frame(x = x, y = borda_ecdf(x) - rand_ecdf(x), Method = "MeanRank",
  stringsAsFactors = F)
bridge_df = rbind(bridge_df, borda_bridge)

toprank_bridge = data.frame(x = x, y = toprank_ecdf(x) - rand_ecdf(x), Method = "TopRank",
  stringsAsFactors = F)
bridge_df = rbind(bridge_df,toprank_bridge)

for(i in 1:length(chea3_libs_ecdf)){
  lib_bridge = data.frame(x = x, y = chea3_libs_ecdf[[i]](x) - rand_ecdf(x), 
    Method = gsub("--"," ", names(chea3_libs_ecdf)[i]),stringsAsFactors = F)
  bridge_df = rbind(bridge_df, lib_bridge)
}

bridge_df = bridge_df[order(bridge_df$Method),]

bridge_plot = ggline(bridge_df, x = "x", y = "y", color = "Method", plot_type = "l", numeric.x.axis = T, xlab = "Scaled Rank", ylab = "D(r) - r")

bridge_plot
```

Report Fraction of Benchmarking Sets reported as Rank 1, Rank 2, etc.
export:10x16
```{r}




generateHist = function(num_benchmarking_sets, max_rank, ranks){
  # breaks = seq(from=10, to= max_rank + 10, by=10)
  quantiles = seq(from = 0.01, to = 1, by=0.01)
  breaks = quantile(seq(1,max_rank,1),quantiles)
  bins = rep(0,length(breaks))
  for(i in 1:length(breaks)){
    if(i==1){
      bins[i] = sum(ranks<=breaks[i])
    }else{
      bins[i] = sum(ranks>breaks[i-1] & ranks<=breaks[i])
    }
  }
  tot_bin = bins/num_benchmarking_sets
  # cum = sapply(breaks, function(x) sum(ranks <= x))
  # frac_cum = cum/num_benchmarking_sets
  sub_bin = bins/length(ranks)
  
  return(rbind(data.frame(quants = quantiles*100, breaks = breaks, bin = tot_bin*100, type = "total"), data.frame(quants = quantiles*100, breaks = breaks, bin = sub_bin*100, type = "subset")))
}
breaks_text = as.character(1:100)
breaks_text[-c(1,25,50,75,100)] = ""

borda_hist = generateHist(num_benchmarking_sets = length(unique(borda$query_set)),
  max_rank = max(borda$rank),
  ranks = borda[borda$class == 1,"rank"]) 

toprank_hist = generateHist(num_benchmarking_sets = length(unique(toprank$query_set)),
  max_rank = max(toprank$rank),
  ranks = toprank[toprank$class == 1,"rank"])

lib_hist = plyr::dlply(chea3_libs,plyr::.(library),function(lib){
  df = generateHist(num_benchmarking_sets = length(unique(lib$query_name)),
  max_rank = max(lib$rank),
  ranks = lib[lib$class == 1,"rank"])
  return(cbind(df,data.frame(library = rep(unique(lib$library),nrow(df)))))
})

borda_hist_plot = ggpubr::ggbarplot(borda_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated MeanRank",position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)

toprank_hist_plot = ggpubr::ggbarplot(toprank_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated TopRank",position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)


lib_hist_plots = lapply(lib_hist, function(lib){

  return(borda_hist_plot = ggpubr::ggbarplot(lib, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = unique(lib$library),position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab")+ scale_x_discrete(labels = breaks_text) )})

hist_plot_list = c(list(meanRank = borda_hist_plot, toprank = toprank_hist_plot), lib_hist_plots)


annotate_figure(ggarrange(plotlist = hist_plot_list,ncol = 1, nrow = length(hist_plot_list), common.legend = T, legend = "top"), left = text_grob("Recovered LOF/GOF TFS (%)",size = 14,rot = 90))



```


Generate table of AUCs and pvalues

```{r}
perform_df = data.frame(Method = c("MeanRank","TopRank",gsub("--"," ",names(chea3_libs_pr))), 
  'ROC AUC' = c(signif(borda_roc$auc,3),
  signif(toprank_roc$auc,3),
  signif(as.numeric(unlist(lapply(chea3_libs_roc,'[[',1))),3)),
  'PR AUC' = c(signif(borda_pr$auc,3),
  signif(toprank_pr$auc,3),
  signif(as.numeric(unlist(lapply(chea3_libs_pr,'[[',1))),3)),
  'Percent Total Recovered' = 
    c(signif(borda_hist[borda_hist$quants == 1 & as.character(borda_hist$type) == "total","bin"],3),
      signif(toprank_hist[toprank_hist$quants == 1 & as.character(toprank_hist$type) == "total","bin"],3), 
      signif(unlist(lapply(lib_hist,function(x){
        return(x[x$quants == 1  & as.character(x$type) == "total","bin"])
      })),3)
      ),
  'Percent Subset Recovered' = 
    c(signif(borda_hist[borda_hist$quants == 1 & as.character(borda_hist$type) == "subset","bin"],3),
      signif(toprank_hist[toprank_hist$quants == 1 & as.character(toprank_hist$type) == "subset","bin"],3), 
      signif(unlist(lapply(lib_hist,function(x){
        return(x[x$quants == 1  & as.character(x$type) == "subset","bin"])
      })),3)),
      
  
  stringsAsFactors = F)
colnames(perform_df) = gsub("\\.", " ", colnames(perform_df))
  
perform_table = ggtexttable(perform_df, rows = NULL, theme = ttheme("minimal"))
perform_table
```


Generate heat map to show transcription factor coverage

```{r}

lit_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ChEA"]]),"_"),"[[",1))))

remap_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ReMap"]]),"_"),"[[",1))))

gtex_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["GTEx"]]),"_"),"[[",1))))

archs4_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ARCHS4"]]),"_"),"[[",1))))

encode_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ENCODE"]]),"_"),"[[",1))))

enrichr_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["Enrichr"]]),"_"),"[[",1))))

perturb_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["Perturbations"]]),"_"),"[[",1))))

integ_tfs = unique(gsub("-","",c(enrichr_tfs, encode_tfs, archs4_tfs, gtex_tfs, remap_tfs, lit_tfs)))

all_tfs = toupper(unique(gsub("-","",na.omit(genesetr::HGNCapproved(chea3::tfs,untranslatable.na = T)))))

heatmap_mat = matrix(0, ncol = length(all_tfs),nrow = 6)

colnames(heatmap_mat) = all_tfs
rownames(heatmap_mat) = c("ARCHS4 Coexpression","ENCODE ChIP-seq","Enrichr Queries","GTEx Coexpression","Literature ChIP-seq","ReMap ChIP-seq")

#populate rows
heatmap_mat["ARCHS4 Coexpression",colnames(heatmap_mat) %in% archs4_tfs] = 1
heatmap_mat["ENCODE ChIP-seq", colnames(heatmap_mat) %in% encode_tfs] = 1
heatmap_mat["Literature ChIP-seq", colnames(heatmap_mat) %in% lit_tfs] = 1
heatmap_mat["ReMap ChIP-seq", colnames(heatmap_mat) %in% remap_tfs] = 1
heatmap_mat["Enrichr Queries", colnames(heatmap_mat) %in% enrichr_tfs] = 1
heatmap_mat["GTEx Coexpression", colnames(heatmap_mat) %in% gtex_tfs] = 1

#cluster
ord = hclust(dist(t(heatmap_mat), method = "euclidean"), method = "ward.D" )$order
ord2 = hclust(dist(heatmap_mat, method = "euclidean"), method = "ward.D")$order


heatmap_melt = reshape2::melt(heatmap_mat)
colnames(heatmap_melt) = c("Library","TF","value")
heatmap_melt$TF = factor(heatmap_melt$TF, levels = colnames(heatmap_mat)[ord])
heatmap_melt$Library = factor(heatmap_melt$Library, levels = rownames(heatmap_mat)[order(rownames(heatmap_mat),decreasing = T)])

tf_heat = ggplot(heatmap_melt, aes(x=TF, y = Library))  + geom_tile(aes(fill = value)) + scale_fill_gradient(low = "white", high = "steelblue") + theme(axis.text.x=element_blank(), legend.position="none", axis.ticks=element_blank())

tf_heat
```

Generate library information plot

```{r}
archs4_interact = genesetr::toLongDF(chea3::ARCHS4)
archs4_interact$a = paste(archs4_interact$set_name,archs4_interact$gene)
archs4_interact = sum(!duplicated(archs4_interact$a))

remap_interact = genesetr::toLongDF(chea3::ReMap)
remap_interact$a = paste(remap_interact$set_name,remap_interact$gene)
remap_interact = sum(!duplicated(remap_interact$a))

lit_interact = genesetr::toLongDF(chea3::ChEA)
lit_interact$a = paste(lit_interact$set_name,lit_interact$gene)
lit_interact = sum(!duplicated(lit_interact$a))


encode_interact = genesetr::toLongDF(chea3::ENCODE)
encode_interact$a = paste(encode_interact$set_name,encode_interact$gene)
encode_interact = sum(!duplicated(encode_interact$a))

gtex_interact = genesetr::toLongDF(chea3::GTEx)
gtex_interact$a = paste(gtex_interact$set_name,gtex_interact$gene)
gtex_interact = sum(!duplicated(gtex_interact$a))

enrichr_interact = genesetr::toLongDF(chea3::Enrichr)
enrichr_interact$a = paste(enrichr_interact$set_name,enrichr_interact$gene)
enrichr_interact = sum(!duplicated(enrichr_interact$a))

libnames = rownames(heatmap_mat)[order()]
lib_info_df = data.frame(Library = c("ARCHS4 Coexpression","ENCODE ChIP-seq","Enrichr Queries","GTEx Coexpression","Literature ChIP-seq","ReMap ChIP-seq"), 
  'Unique TFs' = prettyNum(c(length(archs4_tfs), length(encode_tfs), 
    length(enrichr_tfs), length(gtex_tfs), length(lit_tfs), length(remap_tfs)),big.mark = ",",scientific = F), 
  'TF Interactions' = prettyNum(c(archs4_interact,encode_interact,enrichr_interact,gtex_interact,lit_interact,remap_interact),big.mark = ",",scientific = F), 
  'Gene Sets' = prettyNum(c(length(chea3::ARCHS4),length(chea3::ENCODE),
    length(chea3::Enrichr),length(chea3::GTEx),length(chea3::ChEA),
    length(chea3::ReMap)),big.mark = ",",scientific = F),
  'TF Coverage' = rep('XXXXXXXXXXXXXXXXXXXXXXXXXXXX', 6),stringsAsFactors = F)
colnames(lib_info_df) = gsub("\\."," ",colnames(lib_info_df))

lib_info_table = ggtexttable(lib_info_df, rows = NULL, theme = ttheme("minimal"))
lib_info_table

```

```{r}
perform_df$Method

perform_df$TF_coverage = c(length(integ_tfs), length(integ_tfs), length(archs4_tfs),
  length(encode_tfs), length(enrichr_tfs), length(gtex_tfs), length(lit_tfs), 
  length(remap_tfs))

colnames(perform_df) = gsub(" ","_",colnames(perform_df))

auc_summary_plot = ggpubr::ggscatter(data = perform_df, x = 'ROC_AUC', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "ROC AUC", ylab = "PR AUC", size = 4)
auc_summary_plot

```


Generate Coverage versus Benchmarking Metric Plots

```{r}

roc_coverage_plot = ggpubr::ggscatter(data = perform_df, x = 'TF_coverage', y = 'ROC_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "ROC AUC", size = 4)
pr_coverage_plot = ggpubr::ggscatter(data = perform_df, x = 'TF_coverage', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "PR AUC", size = 4)
subset_coverage_plot = ggpubr::ggscatter(data = perform_df, x = 'TF_coverage', y = 'Percent_Subset_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Subset Recovered", size = 4, ylim = c(0,15))
total_coverage_plot = ggpubr::ggscatter(data = perform_df, x = 'TF_coverage', y = 'Percent_Total_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Total Recovered", size = 4, ylim = c(0,15))




roc_coverage_plot
pr_coverage_plot
subset_coverage_plot
total_coverage_plot


```



Arrange figures
```{r}
library(gridExtra)

ggarrange(auc_summary_plot,bootstrap_roc_plot,bootstrap_pr_plot,bridge_plot,labels = "auto", common.legend = T, legend = "bottom")
ggarrange(tf_heat,subset_coverage_plot,labels = "auto",nrow = 1,ncol = 2)
ggarrange(roc_coverage_plot,pr_coverage_plot,total_coverage_plot, nrow = 1, ncol = 3,labels = "auto")




```


Save workspace

```{r}
#save workspace
 save.image(file = '/Volumes/backup2/ChEA3_API_Benchmark/plot_chea3_benchmarks_pub_version_revisions.Rdata')

```

Load workspace
```{r}
#load workspace
 load(file = '/Volumes/backup2/ChEA3_API_Benchmark/plot_chea3_benchmarks_pub_version_revisions.Rdata')

```


