---
title: "R Notebook"
output: html_notebook
---

Conduct human and mouse benchmarks separately

```{r}

rm(list = ls())

library(ggpubr)


borda = read.table("/Volumes/backup2/ChEA3_API_Benchmark/meanrank_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)

toprank = read.table("/Volumes/backup2/ChEA3_API_Benchmark/toprank_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)

chea3_libs = read.table("/Volumes/backup2/ChEA3_API_Benchmark/library_results_toptfs_trial.tsv", header = T, stringsAsFactors = F)

h_borda = borda[grepl("HUMAN",borda$query_set),]
m_borda = borda[grepl("MOUSE",borda$query_set),]
h_toprank = toprank[grepl("HUMAN",toprank$query_set),]
m_toprank = toprank[grepl("MOUSE",toprank$query_set),]
h_chea3_libs = chea3_libs[grepl("HUMAN",chea3_libs$query_name),]
m_chea3_libs = chea3_libs[grepl("MOUSE",chea3_libs$query_name),]


```

MOUSE


```{r}
library(goftest)

m_borda_ks = ad.test(m_borda[m_borda$class == 1, "scaled_rank"])
m_borda_ks

m_toprank_ks = ad.test(m_toprank[m_toprank$class == 1, "scaled_rank"])
m_toprank_ks

m_chea3_libs_ks = plyr::dlply(m_chea3_libs,plyr::.(library),function(x){
  return(ad.test(x[x$class == 1, "scaled_rank"]))
})

m_chea3_libs_ks

m_borda_ks_df = data.frame(Method = "MeanRank", 
  Pval = signif(-log(m_borda_ks$p.value),4),stringsAsFactors = F)

m_toprank_ks_df = data.frame(Method = "m_toprank", Pval = signif(-log(m_toprank_ks$p.value),4), stringsAsFactors = F)

m_libs_ks_df = plyr::ldply(m_chea3_libs_ks,function(x){
  return(data.frame(Pval = signif(-log(x$p.value),4), stringsAsFactors = F))})





```


Compute bootstrapped ROC Curves

```{r}
library(ggplot2)

n=5000

bootstrap_ROCcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    roc = PRROC::roc.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(roc$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + roc$auc
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(FPR = mean(x[,1]), TPR = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
}

#chea3 m_borda
m_borda_roc = bootstrap_ROCcurve(1-m_borda$scaled_rank[m_borda$class==1],1-m_borda$scaled_rank[m_borda$class==0],n)

#chea3 m_toprank
m_toprank_roc = bootstrap_ROCcurve(1-m_toprank$scaled_rank[m_toprank$class==1],1-m_toprank$scaled_rank[m_toprank$class==0], n)


#chea3 libs
m_chea3_libs_roc = plyr::dlply(m_chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_ROCcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot Bootstrapped ROC curves
```{r}
m_roc = data.frame()

m_roc = rbind(m_roc, data.frame(Method = "MeanRank",
  FPR = m_borda_roc$approx$x, TPR = m_borda_roc$approx$y, stringsAsFactors = F))

m_roc = rbind(m_roc, data.frame(Method = "TopRank",
  FPR = m_toprank_roc$approx$x, TPR = m_toprank_roc$approx$y, stringsAsFactors = F))

m_libs_roc = plyr::ldply(m_chea3_libs_roc,function(lib){
  return(data.frame(
    FPR = lib$approx$x, TPR = lib$approx$y, stringsAsFactors = F))
})

m_libs_roc$Method = gsub("--"," ",m_libs_roc$library)
m_libs_roc$library = NULL

m_roc = rbind(m_roc, m_libs_roc)

m_roc = m_roc[order(m_roc$Method),]
m_bootstrap_roc_plot = ggline(m_roc, x = "FPR", y = "TPR", color = "Method", plot_type = "l",numeric.x.axis = T) 

m_bootstrap_roc_plot

```

Compute bootstrapped PR curves

```{r}

n=5000

bootstrap_PRcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    pr = PRROC::pr.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(pr$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + pr$auc.integral
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(recall = mean(x[,1]), precision = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
 }


#chea3 m_borda
m_borda_pr = bootstrap_PRcurve(1-m_borda$scaled_rank[m_borda$class==1],1-m_borda$scaled_rank[m_borda$class==0],n)

#chea3 m_toprank
m_toprank_pr = bootstrap_PRcurve(1-m_toprank$scaled_rank[m_toprank$class==1],1-m_toprank$scaled_rank[m_toprank$class==0], n)


#chea3 libs
m_chea3_libs_pr = plyr::dlply(m_chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_PRcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot bootstrapped PR curves

```{r}

m_pr = data.frame()

m_pr = rbind(m_pr, data.frame(Method = "MeanRank",
  Recall = m_borda_pr$approx$x, Precision = m_borda_pr$approx$y, stringsAsFactors = F))

m_pr = rbind(m_pr, data.frame(Method = "TopRank",
  Recall = m_toprank_pr$approx$x, Precision = m_toprank_pr$approx$y, stringsAsFactors = F))

m_libs_pr = plyr::ldply(m_chea3_libs_pr,function(lib){
  return(data.frame(
    Recall = lib$approx$x, Precision = lib$approx$y, stringsAsFactors = F))
})

m_libs_pr$Method = gsub("--"," ",m_libs_pr$library)
m_libs_pr$library = NULL

m_pr = rbind(m_pr, m_libs_pr)

m_pr = m_pr[order(m_pr$Method),]
m_bootstrap_pr_plot = ggline(m_pr, x = "Recall", y = "Precision", color = "Method", plot_type = "l",numeric.x.axis = T) 

m_bootstrap_pr_plot

```


Compute and plot bridge plots 
```{r}


m_borda_ecdf = ecdf(m_borda$scaled_rank[m_borda$class == 1])

m_toprank_ecdf = ecdf(m_toprank$scaled_rank[m_toprank$class == 1])

m_chea3_libs_ecdf = plyr::dlply(m_chea3_libs,plyr::.(library),function(lib){
  return(ecdf(lib$scaled_rank[lib$class == 1]))
})

m_rand_ecdf = ecdf(runif(sum(m_borda$class == 1), min = 0, max = 1))

x = seq(0,1,0.001)

m_bridge_df = data.frame()

m_borda_bridge = data.frame(x = x, y = m_borda_ecdf(x) - m_rand_ecdf(x), Method = "MeanRank",
  stringsAsFactors = F)
m_bridge_df = rbind(m_bridge_df, m_borda_bridge)

m_toprank_bridge = data.frame(x = x, y = m_toprank_ecdf(x) - m_rand_ecdf(x), Method = "TopRank",
  stringsAsFactors = F)
m_bridge_df = rbind(m_bridge_df,m_toprank_bridge)

for(i in 1:length(m_chea3_libs_ecdf)){
  m_lib_bridge = data.frame(x = x, y = m_chea3_libs_ecdf[[i]](x) - m_rand_ecdf(x), 
    Method = gsub("--"," ", names(m_chea3_libs_ecdf)[i]),stringsAsFactors = F)
  m_bridge_df = rbind(m_bridge_df, m_lib_bridge)
}

m_bridge_df = m_bridge_df[order(m_bridge_df$Method),]

m_bridge_plot = ggline(m_bridge_df, x = "x", y = "y", color = "Method", plot_type = "l", numeric.x.axis = T, xlab = "Scaled Rank", ylab = "D(r) - r", ylim = c(0,0.5))

m_bridge_plot
```

Report Fraction of Benchmarking Sets reported as Rank 1, Rank 2, etc.
Exported 11x15 in
```{r}

generateHist = function(num_benchmarking_sets, max_rank, ranks){
  # breaks = seq(from=10, to= max_rank + 10, by=10)
  quantiles = seq(from = 0.01, to = 1, by=0.01)
  breaks = quantile(seq(1,max_rank,1),quantiles)
  bins = rep(0,length(breaks))
  for(i in 1:length(breaks)){
    if(i==1){
      bins[i] = sum(ranks<=breaks[i])
    }else{
      bins[i] = sum(ranks>breaks[i-1] & ranks<=breaks[i])
    }
  }
  tot_bin = bins/num_benchmarking_sets
  # cum = sapply(breaks, function(x) sum(ranks <= x))
  # frac_cum = cum/num_benchmarking_sets
  sub_bin = bins/length(ranks)
  
  return(rbind(data.frame(quants = quantiles*100, breaks = breaks, bin = tot_bin*100, type = "total"), data.frame(quants = quantiles*100, breaks = breaks, bin = sub_bin*100, type = "subset")))
}

breaks_text = as.character(1:100)
breaks_text[-c(1,25,50,75,100)]=""

m_borda_hist = generateHist(num_benchmarking_sets = length(unique(m_borda$query_set)),
  max_rank = max(m_borda$rank),
  ranks = m_borda[m_borda$class == 1,"rank"]) 

m_toprank_hist = generateHist(num_benchmarking_sets = length(unique(m_toprank$query_set)),
  max_rank = max(m_toprank$rank),
  ranks = m_toprank[m_toprank$class == 1,"rank"])

m_lib_hist = plyr::dlply(m_chea3_libs,plyr::.(library),function(lib){
  df = generateHist(num_benchmarking_sets = length(unique(lib$query_name)),
  max_rank = max(lib$rank),
  ranks = lib[lib$class == 1,"rank"])
  return(cbind(df,data.frame(library = rep(unique(lib$library),nrow(df)))))
})

m_borda_hist_plot = ggpubr::ggbarplot(m_borda_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated MeanRank",position = position_dodge(0.9), ylim = c(0,18)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)

m_toprank_hist_plot = ggpubr::ggbarplot(m_toprank_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated TopRank",position = position_dodge(0.9), ylim = c(0,18)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)


m_lib_hist_plots = lapply(m_lib_hist, function(lib){

  return(ggpubr::ggbarplot(lib, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = unique(lib$library),position = position_dodge(0.9), ylim = c(0,18)) + rotate_x_text(45) + rremove("ylab")+ scale_x_discrete(labels = breaks_text) )})


m_hist_plot_list = c(list(meanRank = m_borda_hist_plot, m_toprank = m_toprank_hist_plot), m_lib_hist_plots)


annotate_figure(ggarrange(plotlist = m_hist_plot_list,ncol = 1, nrow = length(m_hist_plot_list), common.legend = T, legend = "top"),  left = text_grob('Recovered LOF/GOF TFS (%)',size = 14,rot=90))



```


Generate table of AUCs and pvalues

```{r}
m_perform_df = data.frame(Method = c("MeanRank","TopRank",gsub("--"," ",names(m_chea3_libs_pr))), 
  'ROC AUC' = c(signif(m_borda_roc$auc,3),
  signif(m_toprank_roc$auc,3),
  signif(as.numeric(unlist(lapply(m_chea3_libs_roc,'[[',1))),3)),
  'PR AUC' = c(signif(m_borda_pr$auc,3),
  signif(m_toprank_pr$auc,3),
  signif(as.numeric(unlist(lapply(m_chea3_libs_pr,'[[',1))),3)),
  'Percent Total Recovered' = 
    c(signif(m_borda_hist[m_borda_hist$quants == 1 & as.character(m_borda_hist$type) == "total","bin"],3),
      signif(m_toprank_hist[m_toprank_hist$quants == 1 & as.character(m_toprank_hist$type) == "total","bin"],3), 
      signif(unlist(lapply(m_lib_hist,function(x){
        return(x[x$quants == 1 & x$type == "total","bin"])
      })),3)
      ),
   'Percent Subset Recovered' = 
    c(signif(m_borda_hist[m_borda_hist$quants == 1 & as.character(m_borda_hist$type) == "subset","bin"],3),
      signif(m_toprank_hist[m_toprank_hist$quants == 1 & as.character(m_toprank_hist$type) == "subset","bin"],3), 
      signif(unlist(lapply(m_lib_hist,function(x){
        return(x[x$quants == 1 & as.character(x$type) == "subset","bin"])
      })),3)
      ),
  
  stringsAsFactors = F)
colnames(m_perform_df) = gsub("\\.", " ", colnames(m_perform_df))
  
m_perform_table = ggtexttable(m_perform_df, rows = NULL, theme = ttheme("minimal"))
m_perform_table
```


Generate heat map to show transcription factor coverage

```{r}

lit_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ChEA"]]),"_"),"[[",1))))

remap_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ReMap"]]),"_"),"[[",1))))

gtex_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["GTEx"]]),"_"),"[[",1))))

archs4_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ARCHS4"]]),"_"),"[[",1))))

encode_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["ENCODE"]]),"_"),"[[",1))))

enrichr_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["Enrichr"]]),"_"),"[[",1))))

perturb_tfs = unique(gsub("-","",unlist(lapply(strsplit(names(chea3::libs[["Perturbations"]]),"_"),"[[",1))))

integ_tfs = unique(gsub("-","",c(enrichr_tfs, encode_tfs, archs4_tfs, gtex_tfs, remap_tfs, lit_tfs)))

all_tfs = toupper(unique(gsub("-","",na.omit(genesetr::HGNCapproved(chea3::tfs,untranslatable.na = T)))))

heatmap_mat = matrix(0, ncol = length(all_tfs),nrow = 6)

colnames(heatmap_mat) = all_tfs
rownames(heatmap_mat) = c("ARCHS4 Coexpression","ENCODE ChIP-seq","Enrichr Queries","GTEx Coexpression","Literature ChIP-seq","ReMap ChIP-seq")

#populate rows
heatmap_mat["ARCHS4 Coexpression",colnames(heatmap_mat) %in% archs4_tfs] = 1
heatmap_mat["ENCODE ChIP-seq", colnames(heatmap_mat) %in% encode_tfs] = 1
heatmap_mat["Literature ChIP-seq", colnames(heatmap_mat) %in% lit_tfs] = 1
heatmap_mat["ReMap ChIP-seq", colnames(heatmap_mat) %in% remap_tfs] = 1
heatmap_mat["Enrichr Queries", colnames(heatmap_mat) %in% enrichr_tfs] = 1
heatmap_mat["GTEx Coexpression", colnames(heatmap_mat) %in% gtex_tfs] = 1

#cluster
ord = hclust(dist(t(heatmap_mat), method = "euclidean"), method = "ward.D" )$order
ord2 = hclust(dist(heatmap_mat, method = "euclidean"), method = "ward.D")$order


heatmap_melt = reshape2::melt(heatmap_mat)
colnames(heatmap_melt) = c("Library","TF","value")
heatmap_melt$TF = factor(heatmap_melt$TF, levels = colnames(heatmap_mat)[ord])
heatmap_melt$Library = factor(heatmap_melt$Library, levels = rownames(heatmap_mat)[order(rownames(heatmap_mat),decreasing = T)])

tf_heat = ggplot(heatmap_melt, aes(x=TF, y = Library))  + geom_tile(aes(fill = value)) + scale_fill_gradient(low = "white", high = "steelblue") + theme(axis.text.x=element_blank(), legend.position="none", axis.ticks=element_blank())

tf_heat
```

Generate library information plot

```{r}
archs4_interact = genesetr::toLongDF(chea3::ARCHS4)
archs4_interact$a = paste(archs4_interact$set_name,archs4_interact$gene)
archs4_interact = sum(!duplicated(archs4_interact$a))

remap_interact = genesetr::toLongDF(chea3::ReMap)
remap_interact$a = paste(remap_interact$set_name,remap_interact$gene)
remap_interact = sum(!duplicated(remap_interact$a))

lit_interact = genesetr::toLongDF(chea3::ChEA)
lit_interact$a = paste(lit_interact$set_name,lit_interact$gene)
lit_interact = sum(!duplicated(lit_interact$a))


encode_interact = genesetr::toLongDF(chea3::ENCODE)
encode_interact$a = paste(encode_interact$set_name,encode_interact$gene)
encode_interact = sum(!duplicated(encode_interact$a))

gtex_interact = genesetr::toLongDF(chea3::GTEx)
gtex_interact$a = paste(gtex_interact$set_name,gtex_interact$gene)
gtex_interact = sum(!duplicated(gtex_interact$a))

enrichr_interact = genesetr::toLongDF(chea3::Enrichr)
enrichr_interact$a = paste(enrichr_interact$set_name,enrichr_interact$gene)
enrichr_interact = sum(!duplicated(enrichr_interact$a))

libnames = rownames(heatmap_mat)[order()]
lib_info_df = data.frame(Library = c("ARCHS4 Coexpression","ENCODE ChIP-seq","Enrichr Queries","GTEx Coexpression","Literature ChIP-seq","ReMap ChIP-seq"), 
  'Unique TFs' = prettyNum(c(length(archs4_tfs), length(encode_tfs), 
    length(enrichr_tfs), length(gtex_tfs), length(lit_tfs), length(remap_tfs)),big.mark = ",",scientific = F), 
  'TF Interactions' = prettyNum(c(archs4_interact,encode_interact,enrichr_interact,gtex_interact,lit_interact,remap_interact),big.mark = ",",scientific = F), 
  'Gene Sets' = prettyNum(c(length(chea3::ARCHS4),length(chea3::ENCODE),
    length(chea3::Enrichr),length(chea3::GTEx),length(chea3::ChEA),
    length(chea3::ReMap)),big.mark = ",",scientific = F),
  'TF Coverage' = rep('XXXXXXXXXXXXXXXXXXXXXXXXXXXX', 6),stringsAsFactors = F)
colnames(lib_info_df) = gsub("\\."," ",colnames(lib_info_df))

lib_info_table = ggtexttable(lib_info_df, rows = NULL, theme = ttheme("minimal"))
lib_info_table

```

```{r}
m_perform_df$Method

m_perform_df$TF_coverage = c(length(integ_tfs), length(integ_tfs), length(archs4_tfs),
  length(encode_tfs), length(enrichr_tfs), length(gtex_tfs), length(lit_tfs), 
  length(remap_tfs))

colnames(m_perform_df) = gsub(" ","_",colnames(m_perform_df))

m_auc_summary_plot = ggpubr::ggscatter(data = m_perform_df, x = 'ROC_AUC', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "ROC AUC", ylab = "PR AUC", size = 4)
m_auc_summary_plot

```


Generate Coverage versus Benchmarking Metric Plots

```{r}

m_roc_coverage_plot = ggpubr::ggscatter(data = m_perform_df, x = 'TF_coverage', y = 'ROC_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "ROC AUC", size = 4)
m_pr_coverage_plot = ggpubr::ggscatter(data = m_perform_df, x = 'TF_coverage', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "PR AUC", size = 4)
m_subset_coverage_plot = ggpubr::ggscatter(data = m_perform_df, x = 'TF_coverage', y = 'Percent_Subset_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Subset Recovered", size = 4, ylim = c(0,18))
m_tot_coverage_plot = ggpubr::ggscatter(data = m_perform_df, x = 'TF_coverage', y = 'Percent_Total_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Total Recovered", size = 4, ylim = c(0,18))


m_roc_coverage_plot
m_pr_coverage_plot
m_tot_coverage_plot
m_subset_coverage_plot
```

HUMAN

```{r}
library(goftest)

h_borda_ks = ad.test(h_borda[h_borda$class == 1, "scaled_rank"])
h_borda_ks

h_toprank_ks = ad.test(h_toprank[h_toprank$class == 1, "scaled_rank"])
h_toprank_ks

h_chea3_libs_ks = plyr::dlply(h_chea3_libs,plyr::.(library),function(x){
  return(ad.test(x[x$class == 1, "scaled_rank"]))
})

h_chea3_libs_ks

h_borda_ks_df = data.frame(Method = "MeanRank", 
  Pval = signif(-log(h_borda_ks$p.value),4),stringsAsFactors = F)

h_toprank_ks_df = data.frame(Method = "TopRank", Pval = signif(-log(h_toprank_ks$p.value),4), stringsAsFactors = F)

h_libs_ks_df = plyr::ldply(h_chea3_libs_ks,function(x){
  return(data.frame(Pval = signif(-log(x$p.value),4), stringsAsFactors = F))})





```


Compute bootstrapped ROC Curves

```{r}
library(ggplot2)

n=5000

bootstrap_ROCcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    roc = PRROC::roc.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(roc$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + roc$auc
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(FPR = mean(x[,1]), TPR = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
}

#chea3 h_borda
h_borda_roc = bootstrap_ROCcurve(1-h_borda$scaled_rank[h_borda$class==1],1-h_borda$scaled_rank[h_borda$class==0],n)

#chea3 h_toprank
h_toprank_roc = bootstrap_ROCcurve(1-h_toprank$scaled_rank[h_toprank$class==1],1-h_toprank$scaled_rank[h_toprank$class==0], n)


#chea3 libs
h_chea3_libs_roc = plyr::dlply(h_chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_ROCcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot Bootstrapped ROC curves
```{r}
h_roc = data.frame()

h_roc = rbind(h_roc, data.frame(Method = "MeanRank",
  FPR = h_borda_roc$approx$x, TPR = h_borda_roc$approx$y, stringsAsFactors = F))

h_roc = rbind(h_roc, data.frame(Method = "TopRank",
  FPR = h_toprank_roc$approx$x, TPR = h_toprank_roc$approx$y, stringsAsFactors = F))

h_libs_roc = plyr::ldply(h_chea3_libs_roc,function(lib){
  return(data.frame(
    FPR = lib$approx$x, TPR = lib$approx$y, stringsAsFactors = F))
})

h_libs_roc$Method = gsub("--"," ",h_libs_roc$library)
h_libs_roc$library = NULL

h_roc = rbind(h_roc, h_libs_roc)

h_roc = h_roc[order(h_roc$Method),]
h_bootstrap_roc_plot = ggline(h_roc, x = "FPR", y = "TPR", color = "Method", plot_type = "l",numeric.x.axis = T) 

h_bootstrap_roc_plot

```

Compute bootstrapped PR curves

```{r}

n=5000

bootstrap_PRcurve = function(ones, zeros, n){
  size_group1 = length(ones)
  sum_AUC = 0
  curves = list()
  for(i in 1:n){
    pr = PRROC::pr.curve(ones, sample(zeros,size_group1), curve = T)
    curve = as.data.frame(pr$curve)
    curve$bin = .bincode(curve[,3],seq(0,1,0.0005))
    sum_AUC = sum_AUC + pr$auc.integral
    curves[[i]] = curve
  }
  curves = do.call(rbind,curves)
  auc = sum_AUC/n
  approx = approx(curves[,1],curves[,2],n=100)
  c = plyr::ddply(curves,plyr::.(bin),function(x){
    return(data.frame(recall = mean(x[,1]), precision = mean(x[,2])))
  })
  return(list(auc = auc, curve = c[,2:3], approx = approx))
 }


#chea3 h_borda
h_borda_pr = bootstrap_PRcurve(1-h_borda$scaled_rank[h_borda$class==1],1-h_borda$scaled_rank[h_borda$class==0],n)

#chea3 h_toprank
h_toprank_pr = bootstrap_PRcurve(1-h_toprank$scaled_rank[h_toprank$class==1],1-h_toprank$scaled_rank[h_toprank$class==0], n)


#chea3 libs
h_chea3_libs_pr = plyr::dlply(h_chea3_libs,plyr::.(library),function(lib){
  return(bootstrap_PRcurve(1-lib$scaled_rank[lib$class==1],1-lib$scaled_rank[lib$class==0],n))
})


```

Plot bootstrapped PR curves

```{r}

h_pr = data.frame()

h_pr = rbind(h_pr, data.frame(Method = "MeanRank",
  Recall = h_borda_pr$approx$x, Precision = h_borda_pr$approx$y, stringsAsFactors = F))

h_pr = rbind(h_pr, data.frame(Method = "TopRank",
  Recall = h_toprank_pr$approx$x, Precision = h_toprank_pr$approx$y, stringsAsFactors = F))

h_libs_pr = plyr::ldply(h_chea3_libs_pr,function(lib){
  return(data.frame(
    Recall = lib$approx$x, Precision = lib$approx$y, stringsAsFactors = F))
})

h_libs_pr$Method = gsub("--"," ",h_libs_pr$library)
h_libs_pr$library = NULL

h_pr = rbind(h_pr, h_libs_pr)

h_pr = h_pr[order(h_pr$Method),]
h_bootstrap_pr_plot = ggline(h_pr, x = "Recall", y = "Precision", color = "Method", plot_type = "l",numeric.x.axis = T) 

h_bootstrap_pr_plot

```


Compute and plot bridge plots 
```{r}


h_borda_ecdf = ecdf(h_borda$scaled_rank[h_borda$class == 1])

h_toprank_ecdf = ecdf(h_toprank$scaled_rank[h_toprank$class == 1])

h_chea3_libs_ecdf = plyr::dlply(h_chea3_libs,plyr::.(library),function(lib){
  return(ecdf(lib$scaled_rank[lib$class == 1]))
})

h_rand_ecdf = ecdf(runif(sum(h_borda$class == 1), min = 0, max = 1))

x = seq(0,1,0.001)

h_bridge_df = data.frame()

h_borda_bridge = data.frame(x = x, y = h_borda_ecdf(x) - h_rand_ecdf(x), Method = "MeanRank",
  stringsAsFactors = F)
h_bridge_df = rbind(h_bridge_df, h_borda_bridge)

h_toprank_bridge = data.frame(x = x, y = h_toprank_ecdf(x) - h_rand_ecdf(x), Method = "TopRank",
  stringsAsFactors = F)
h_bridge_df = rbind(h_bridge_df,h_toprank_bridge)

for(i in 1:length(h_chea3_libs_ecdf)){
  h_lib_bridge = data.frame(x = x, y = h_chea3_libs_ecdf[[i]](x) - h_rand_ecdf(x), 
    Method = gsub("--"," ", names(h_chea3_libs_ecdf)[i]),stringsAsFactors = F)
  h_bridge_df = rbind(h_bridge_df, h_lib_bridge)
}

h_bridge_df = h_bridge_df[order(h_bridge_df$Method),]

h_bridge_plot = ggline(h_bridge_df, x = "x", y = "y", color = "Method", plot_type = "l", numeric.x.axis = T, xlab = "Scaled Rank", ylab = "D(r) - r", ylim = c(0,0.5))

h_bridge_plot
```

Report Fraction of Benchmarking Sets reported as Rank 1, Rank 2, etc.
Exported 11x15 inches
```{r}


generateHist = function(num_benchmarking_sets, max_rank, ranks){
  # breaks = seq(from=10, to= max_rank + 10, by=10)
  quantiles = seq(from = 0.01, to = 1, by=0.01)
  breaks = quantile(seq(1,max_rank,1),quantiles)
  bins = rep(0,length(breaks))
  for(i in 1:length(breaks)){
    if(i==1){
      bins[i] = sum(ranks<=breaks[i])
    }else{
      bins[i] = sum(ranks>breaks[i-1] & ranks<=breaks[i])
    }
  }
  tot_bin = bins/num_benchmarking_sets
  # cum = sapply(breaks, function(x) sum(ranks <= x))
  # frac_cum = cum/num_benchmarking_sets
  sub_bin = bins/length(ranks)
  
  return(rbind(data.frame(quants = quantiles*100, breaks = breaks, bin = tot_bin*100, type = "total"), data.frame(quants = quantiles*100, breaks = breaks, bin = sub_bin*100, type = "subset")))
}

breaks_text = as.character(1:100)
breaks_text[-c(1,25,50,75,100)] = ""

h_borda_hist = generateHist(num_benchmarking_sets = length(unique(h_borda$query_set)),
  max_rank = max(h_borda$rank),
  ranks = h_borda[h_borda$class == 1,"rank"]) 

h_toprank_hist = generateHist(num_benchmarking_sets = length(unique(h_toprank$query_set)),
  max_rank = max(h_toprank$rank),
  ranks = h_toprank[h_toprank$class == 1,"rank"])

h_lib_hist = plyr::dlply(h_chea3_libs,plyr::.(library),function(lib){
  df = generateHist(num_benchmarking_sets = length(unique(lib$query_name)),
  max_rank = max(lib$rank),
  ranks = lib[lib$class == 1,"rank"])
  return(cbind(df,data.frame(library = rep(unique(lib$library),nrow(df)))))
})

h_borda_hist_plot = ggpubr::ggbarplot(h_borda_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated MeanRank",position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)

h_toprank_hist_plot = ggpubr::ggbarplot(m_toprank_hist, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = "Integrated TopRank",position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab") + scale_x_discrete(labels = breaks_text)


h_lib_hist_plots = lapply(h_lib_hist, function(lib){

  return(ggpubr::ggbarplot(lib, x = "quants", y = "bin", xlab = "Rank Percentile", ylab = "Recovered LOF/GOF TFs (%)", color = "type", title = unique(lib$library),position = position_dodge(0.9), ylim = c(0,15)) + rotate_x_text(45) + rremove("ylab")+ scale_x_discrete(labels = breaks_text) )})


h_hist_plot_list = c(list(meanRank = h_borda_hist_plot, h_toprank = h_toprank_hist_plot), h_lib_hist_plots)


annotate_figure(ggarrange(plotlist = h_hist_plot_list,ncol = 1, nrow = length(h_hist_plot_list), common.legend = T, legend = "top"),  left = text_grob('Recovered LOF/GOF TFS (%)',size = 14,rot=90))


```


Generate table of AUCs and pvalues

```{r}
h_perform_df = data.frame(Method = c("MeanRank","TopRank",gsub("--"," ",names(h_chea3_libs_pr))), 
  'ROC AUC' = c(signif(h_borda_roc$auc,3),
  signif(h_toprank_roc$auc,3),
  signif(as.numeric(unlist(lapply(h_chea3_libs_roc,'[[',1))),3)),
  'PR AUC' = c(signif(h_borda_pr$auc,3),
  signif(h_toprank_pr$auc,3),
  signif(as.numeric(unlist(lapply(h_chea3_libs_pr,'[[',1))),3)),
  'Percent Total Recovered' = 
    c(signif(h_borda_hist[h_borda_hist$quant == 1 & h_borda_hist$type == "total","bin"],3),
      signif(h_toprank_hist[h_toprank_hist$quant == 1 & h_toprank_hist$type == "total","bin"],3), 
      signif(unlist(lapply(h_lib_hist,function(x){
        return(x[x$quant == 1 & x$type == "total","bin"])
      })),3)
      ),
  'Percent Subset Recovered' = 
    c(signif(h_borda_hist[h_borda_hist$quant == 1 & h_borda_hist$type == "subset","bin"],3),
      signif(h_toprank_hist[h_toprank_hist$quant == 1 & h_toprank_hist$type == "subset","bin"],3), 
      signif(unlist(lapply(h_lib_hist,function(x){
        return(x[x$quant == 1 & x$type == "subset","bin"])
      })),3)
      ),
  
  stringsAsFactors = F)
colnames(h_perform_df) = gsub("\\.", " ", colnames(h_perform_df))
  
h_perform_table = ggtexttable(h_perform_df, rows = NULL, theme = ttheme("minimal"))
h_perform_table
```


```{r}
h_perform_df$Method

h_perform_df$TF_coverage = c(length(integ_tfs), length(integ_tfs), length(archs4_tfs),
  length(encode_tfs), length(enrichr_tfs), length(gtex_tfs), length(lit_tfs), 
  length(remap_tfs))

colnames(h_perform_df) = gsub(" ","_",colnames(h_perform_df))

h_auc_summary_plot = ggpubr::ggscatter(data = h_perform_df, x = 'ROC_AUC', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "ROC AUC", ylab = "PR AUC", size = 4)
h_auc_summary_plot

```


Generate Coverage versus Benchmarking Metric Plots

```{r}

h_roc_coverage_plot = ggpubr::ggscatter(data = h_perform_df, x = 'TF_coverage', y = 'ROC_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "ROC AUC", size = 4)
h_pr_coverage_plot = ggpubr::ggscatter(data = h_perform_df, x = 'TF_coverage', y = 'PR_AUC', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "PR AUC", size = 4)
h_tot_coverage_plot = ggpubr::ggscatter(data = h_perform_df, x = 'TF_coverage', y = 'Percent_Total_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Total Recovered", size = 4, ylim = c(0,18))
h_sub_coverage_plot = ggpubr::ggscatter(data = h_perform_df, x = 'TF_coverage', y = 'Percent_Subset_Recovered', label = 'Method', repel = T, label.rectangle = F, palette = "jco",  alpha = 0.25, xlab = "TF Coverage", ylab = "Percent Subset Recovered", size = 4, ylim = c(0,18))


h_roc_coverage_plot
h_pr_coverage_plot
h_tot_coverage_plot
h_sub_coverage_plot

```


Arrange Mouse and Human Plots

Export size: plot1 ?
plot2 
```{r}
ggarrange(h_bootstrap_roc_plot,m_bootstrap_roc_plot,h_bootstrap_pr_plot,m_bootstrap_pr_plot,h_bridge_plot,m_bridge_plot,
ncol = 2, nrow = 3, common.legend = T, legend = "bottom", labels = "auto") + theme(plot.margin = unit(c(2,1,1,1),"cm"))

ggarrange(h_auc_summary_plot, m_auc_summary_plot, h_roc_coverage_plot, m_roc_coverage_plot, h_pr_coverage_plot, m_pr_coverage_plot, h_tot_coverage_plot, m_tot_coverage_plot, h_sub_coverage_plot, m_subset_coverage_plot, ncol = 2, nrow = 5, labels = "auto") + theme(plot.margin = unit(c(2,1,1,1),"cm"))
```

Save 
```{r}
save.image(file = "/volumes/Backup2/ChEA3_API_Benchmark/mouse_human_benchmarking.Rdata")

```



